Wigglyhedra

### 2024-10-12 FPSAC25/wigglyhedraFPSAC.tex
minor changes

%% put your French abstract here, or comment this out if you don't have one
\resume{Motivés par la théorie catégorique des représentations, nous étudions le slalomèdre. Son complexe de bord encode les dissections slalomantes (ensembles deux-à-deux pointés et sans croisement d'arcs slalomant autour de $n+2$ points sur une ligne). Son graphe est le diagramme de Hasse du sous-treillis de l'ordre faible induit par les permutations slalomantes (permutations de $[2n]$ évitant~$2j-1 \cdots i \cdots 2j$ pour~$i < 2j-1$ et~$2j \cdots k \cdots 2j-1$ pour~$k > 2j$). Il contient des copies de tous les treillis Cambriens et associaèdres Cambriens de type~$A_n$.}

@@

Motivated by categorical representation theory, we consider the wiggly complex (\cref{sec:wigglyPseudotriangulations}), whose vertices are arcs wiggling around $n+2$ points on a line, and whose faces are sets of pairwise pointed and non-crossing wiggly arcs.
It is a $(2n-1)$-dimensional manifold whose maximal faces are wiggly pseudotriangulations of a disc.
We show (\cref{sec:bijection}) that the flip graph on wiggly pseudotriangulations is isomorphic to the cover graph of the sublattice of the weak order induced by wiggly permutations (\cref{sec:wigglyPermutations}), the permutations of $[2n]$ avoiding~$2j-1 \cdots i \cdots 2j$ for~$i < 2j-1$ and~$2j \cdots k \cdots 2j-1$ for~$k > 2j$.
Inspired by similar constructions for finite type cluster complexes~\cite{HohlwegLangeThomas,HohlwegPilaudStella}, and for support $\tau$-tilting complexes on gentle algebras~\cite{PaluPilaudPlamondon-nonkissing}, we then provide geometric realizations of the wiggly complex.
We first construct the wiggly fan (\cref{sec:wigglyFan}), a complete simplicial fan realizing the wiggly complex and supported by the $\b{g}$-vectors of the wiggly arcs.
We then define the wigglyhedron (\cref{sec:wigglyhedron}), a polytope whose normal fan is the wiggly fan and whose support function is given by the incompatibility numbers of wiggly arcs.
Its graph can be oriented to obtain the Hasse diagram of the wiggly lattice, and it contains copies of all type~$A_n$ Cambrian lattice as faces (\cref{sec:CambrianConsiderations}).
To conclude, we conjecture that the wiggly complex of any planar point set is polytopal (\cref{sec:planarPointSets}), and we explore representation theoretic aspects of the wiggly complex (\cref{sec:representationTheory}).

@@

\begin{definition}
\label{def:wigglyComplex}
The \defn{wiggly complex}~$\wigglyComplex_n$ is the simplicial complex of pairwise pointed and non-crossing subsets of internal wiggly arcs.
%In other words, it is the clique complex of the compatibility graph on internal wiggly arcs.

>>>>

%% put your French abstract here, or comment this out if you don't have one
\resume{Motivés par la théorie catégorique des représentations, nous étudions \linebreak l'onduloèdre. Son complexe de bord encode les dissections ondulantes (ensembles d'arcs ondulant autour de $n+2$ points sur une ligne, deux-à-deux pointés et sans croisement). Son graphe est le diagramme de Hasse du sous-treillis de l'ordre faible induit par les permutations ondulantes (permutations de $[2n]$ évitant~$2j-1 \cdots i \cdots 2j$ pour~$i < 2j-1$ et~$2j \cdots k \cdots 2j-1$ pour~$k > 2j$). Il contient des copies de tous les treillis Cambriens et associaèdres Cambriens de type~$A_n$.}

@@

Motivated by categorical representation theory, we consider the wiggly complex (\cref{sec:wigglyPseudotriangulations}), whose vertices are arcs wiggling around $n+2$ points on a line, and whose faces are sets of pairwise pointed and non-crossing wiggly arcs.
It is a $(2n-1)$-dimensional manifold whose maximal faces are wiggly pseudotriangulations of a disc.
We show (\cref{sec:bijection}) that the flip graph on wiggly pseudotriangulations is isomorphic to the cover graph of the sublattice of the weak order induced by wiggly permutations (\cref{sec:wigglyPermutations}), the permutations of $[2n]$ avoiding~$2j-1 \cdots i \cdots 2j$ for~$i < 2j-1$ and~$2j \cdots k \cdots 2j-1$ for~$k > 2j$.
Inspired by similar constructions for finite type cluster complexes~\cite{HohlwegLangeThomas,HohlwegPilaudStella}, and for support $\tau$-tilting complexes on gentle algebras~\cite{PaluPilaudPlamondon-nonkissing}, we then provide geometric realizations of the wiggly complex.
We first construct the wiggly fan (\cref{sec:wigglyFan}), a complete simplicial fan realizing the wiggly complex and supported by the $\b{g}$-vectors of the wiggly arcs.
We then define the wigglyhedron (\cref{sec:wigglyhedron}), a polytope whose normal fan is the wiggly fan and whose support function is given by the incompatibility numbers of wiggly arcs.
Its graph can be oriented to obtain the Hasse diagram of the wiggly lattice, and it contains copies of all type~$A_n$ Cambrian lattices~\cite{Reading-CambrianLattices} as faces (\cref{sec:CambrianConsiderations}).
To conclude, we conjecture that the wiggly complex of any planar point set is polytopal (\cref{sec:planarPointSets}), and we explore representation theoretic aspects of the wiggly complex (\cref{sec:representationTheory}).

@@

\begin{definition}
\label{def:wigglyComplex}
The \defn{wiggly complex}~$\wigglyComplex_n$ is the simplicial complex of subsets of pairwise compatible internal wiggly arcs.
%In other words, it is the clique complex of the compatibility graph on internal wiggly arcs.
